Data Specialization for a Staged Lambda Calculus

### 2014-11-07 paper/splitting.tex
Small refinements to the text in partial value masking.  Verbosely using the terminology stage 1 masking and stage 2 masking.

Given a term in the source language \lang\ where terms of
stage~\bbone\ and stage~\bbtwo\ may be interleaved, \emph{(stage)
  splitting} separates the stage~\bbone\ and stage~\bbtwo\
subcomputations into two separate terms, each performing the
subcomputations corresponding to each stage.  More precisely, given a
\lang\ term~$e$, splitting yields two terms $e_1$ and $e_2$ in a
unstaged target language \langmono\ such that~$e_1$ performs all the
stage~\bbone\ subcomputations in~$e$ and $e_2$~performs all of the
stage~\bbtwo\ subcomputations. Splitting preserves correctness in the sense that evaluation of~$e$ in
the staged language produces the same result as evaluating~$e_1$ and 
then~$e_2$ with the result from~$e_1$.

@@

Recall that \lang\ partial values and residual tables together may
contain content at \emph{both} stages. For example the term:
\begin{lstlisting}
1`(1+2, next{`2`hold{`1`3+4`2`} + 5`1`})`
\end{lstlisting}
evaluates via $\redonesym$ to:
\begin{lstlisting}
2`[yhat|->7+5]` 1`(3,next{`2`yhat`1`})`
\end{lstlisting}
Intuitively, splitting this term should generate an $e1$ that reduces to the stage 1 content of the partial value (3), and an $e2$ corresponding to the stage~\bbtwo\
computation $\letin \yhat {\mathtt{7+5}} \yhat$.
To precisely delineate the stage~\bbone\ and stage~\bbtwo\ components of the partial value, we define 
\emph{partial-value masking} (or simply {\em masking}) functions, written $\masko{{-}}$ and $\maskt{{-};{-}}$, 
which take a residual table and partial value as input and emit a term in \langmono\ that corresponds to only the content belonging to a single stage. \ref{fig:valMask} defines these
masking functions.
% written $\masko{{-}}$ and $\maskt{{-};{-}}$.

@@

\emph{Stage~\bbtwo\ masking} of the partial value $v$ and residual context $\xi$ (written $\maskt{\xi;v}$) produces a term in \langmono\ containing
exactly the stage~\bbtwo\ content of $[\xi;v]$. Stage~\bbtwo\ masking 
requires access to the residual table $\xi$ since all stage~\bbtwo\
content in $v$ has been lifted out into $\xi$. Stage~\bbtwo\ masking discards
stage~\bbone\ content in $v$ by replacing base constants with~\texttt{()},
replacing $\next~\hat{y}$ with~$\hat{y}$, and replacing all $n$-ary term
constructors with an $n$-tuple of masked subterms. Like stage~\bbone\ splitting, stage~\bbtwo\
splititing masks functions by splitting their bodies as general stage~\bbone\ terms, and
returning the stage~\bbtwo\ component. After discarding stage~\bbone\ content from $v$, stage~\bbtwo\ reifies $\xi$ around the result,
producing a term in \langmono.

>>>>

Given a term in the source language \lang\ where terms of
stage~\bbone\ and stage~\bbtwo\ may be interleaved, \emph{(stage)
  splitting} separates the stage~\bbone\ and stage~\bbtwo\
subcomputations into two separate terms, each performing the
subcomputations corresponding to each stage.  More precisely, given a
\lang\ term~$e$, splitting yields two terms $e_1$ and $e_2$ in a
unstaged target language \langmono\ such that~$e_1$ performs all the
stage~\bbone\ subcomputations in~$e$ and $e_2$~performs all of the
stage~\bbtwo\ subcomputations. Splitting preserves correctness in the sense that evaluation of~$e$ in
the staged language produces the same result as evaluating~$e_1$, followed by evaluating 
$e_2$ with the result from~$e_1$.

@@

Recall that \lang\ partial values and residual tables together may
contain content at \emph{both} stages. For example the term:
\begin{lstlisting}
1`(1+2, next{`2`hold{`1`3+4`2`} + 5`1`})`
\end{lstlisting}
evaluates via $\redonesym$ to:
\begin{lstlisting}
2`[yhat|->7+5]` 1`(3,next{`2`yhat`1`})`
\end{lstlisting}
Intuitively, splitting this term should generate an $e1$ that reduces to the stage 1 content of the partial value (3), and an $e2$ corresponding to the stage~\bbtwo\
computation $\letin \yhat {\mathtt{7+5}} \yhat$.
To precisely define the stage~\bbone\ and stage~\bbtwo\ components of the partial value, we define 
\emph{partial-value masking} (or simply {\em masking}) functions, written $\masko{{-}}$ and $\maskt{{-};{-}}$, 
which take a residual table and partial value as input and emit a term in \langmono\ that corresponds to only the content belonging to a single stage. \ref{fig:valMask} defines these
masking functions.
% written $\masko{{-}}$ and $\maskt{{-};{-}}$.

@@

\emph{Stage~\bbtwo\ masking} of the partial value $v$ and residual context $\xi$ (written $\maskt{\xi;v}$) produces a term in \langmono\ containing
exactly the stage~\bbtwo\ content of $[\xi;v]$. Unlike stage~\bbone\ masking, stage~\bbtwo\ masking 
requires access to the residual table $\xi$ because all stage~\bbtwo\
content in $v$ has been lifted out into $\xi$. Stage~\bbtwo\ masking discards
stage~\bbone\ content in $v$ by replacing base constants with~\texttt{()},
replacing $\next~\hat{y}$ with~$\hat{y}$, and replacing all $n$-ary term
constructors with an $n$-tuple of masked subterms. Like stage~\bbone\ splitting, stage~\bbtwo\
splititing masks functions by splitting their bodies as general stage~\bbone\ terms, and
returning the stage~\bbtwo\ component. After discarding stage~\bbone\ content from $v$, stage~\bbtwo\ masking reifies $\xi$ around the result,
producing a term in \langmono.
